---
title: "R_project: Programmation et traitement statistique des données"
output: html_document
---
# Question 1: Estimateurs Non Paramétriques avec récursivité de la dérivée d'une fonction de régression non paramétrique

```{r}
N <- 10000 # sample size
alpha <- 0.3 # selected by cross validation (to check)
e_N <- function(x) {
  return((1/x^alpha))
}
h <- sapply(seq(1, N, 1), e_N) # bandwidth

# K values
# gaussian -> pnorm(x)
# quadratic -> ifelse(abs(x) <= 1, (15/16)*(1-x^2)^2), 0)
# epanechnikov -> ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)

# K prime values
# gaussian -> -(x*(1/sqrt(2*pi))*(exp(-(1/2)*x^2)
# quadratic -> ifelse(abs(x) <= 1, -(15/8)*(x)*(1-x^2), 0)
# epanechnikov -> ifelse(abs(x) <= 1, -(3/2)*(x)*(1-x^2), 0)

# estimator n°1 toolbox

h_N <- function(x) {
  return((1/N)*(sum((Y_N/h)*K((x-X_N)/h))))
}

g_N <- function(x) {
  return((1/N)*(sum((1/h)*K((x-X_N)/h))))
}

h_N_prime <- function(x) {
  return((1/N)*(sum((Y_N/h^2)*K_prime((x-X_N)/h))))
}

g_N_prime <- function(x) {
  return((1/N)*(sum((1/h^2)*K_prime((x-X_N)/h))))
}

es_n1 <- function(x) {
  return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
}

# estimator n°2



es_n2 <- function(x) {
  return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
}

# estimator n°3

es_n3 <- function(x) {
  return((1/N)*(sum(Y_N/(g(x)*h^2)*K_prime((x-X_N)/h))))
}

```

# Question 2: Simulations

## Bandwidth Selection with Cross-Validation

Alpha would depend on the chosen kernel (among gaussian, epanechnikov, quadratic) and on the chosen non parametric estimator.

Here we take K(x) from Epanechnikov and we evaluate alpha depending on the chosen estimator (among Nadaraya Watson (original, Johnston, Wand)).

```{r}
# global scope functions
f <- function(x) {
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) {
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}
g <- function(x) {
  return(1)
}
g_prime <- function(x){
  return(0)
}
K <- function(x) {
  return(ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)) 
}
K_prime <- function(x) {
  return(ifelse(abs(x) <= 1, -(3/2)*x, 0))
}
# data generation
N <- 1000
eps_N <- rnorm(N, 0, 1)
X_N <- runif(N, 0, 1) # p = 1
Y_N <- f(X_N) + eps_N

CV <- function(alpha) { # CV for each estimator given K, g, f, X, Y
  res <- rep(NA, 3)
  X <- X_N
  Y <- Y_N
  for (j in 1:3) {
    cv_score <- 0
    e_N <- function(x) { # make alpha dependent to the estimator
    return((1/x^alpha[j]))
    }
    h <- (1:N)^(-alpha[j])
    H <- h
    for (k in 1:length(X_N)) { # compute CV_score
      X_N <- X
      Y_N <- Y
      h <- H[-k]
      X_N <- X_N[-k]
      Y_N <- Y_N[-k]
      es_n1 <- function(x) {
        return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
      }
      es_n2 <- function(x) {
        return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
      }
      es_n3 <- function(x) {
        return((1/N)*(sum(Y_N/(g(x)*h^2)*K_prime((x-X_N)/h))))
      }
      h_N <- function(x) {
        return((1/N)*(sum((Y_N/h)*K((x-X_N)/h))))
      }
      g_N <- function(x) {
        return((1/N)*(sum((1/h)*K((x-X_N)/h))))
      }
      h_N_prime <- function(x) {
        return((1/N)*(sum((Y_N/h^2)*K_prime((x-X_N)/h))))
      }
      g_N_prime <- function(x) {
        return((1/N)*(sum((1/h^2)*K_prime((x-X_N)/h))))
      }
      if (j == 1) {
        cv_score <- cv_score + (1/N)*(es_n1(X[k]) - f_prime(X[k]))^2
      } else if (j == 2) {
        cv_score <- cv_score + (1/N)*(es_n2(X[k]) - f_prime(X[k]))^2
      } else {
        cv_score <- cv_score + (1/N)*(es_n3(X[k]) - f_prime(X[k]))^2
      }
    }
    res[j] <- cv_score
  }
  return(sum(res))
}

# Optimal bandwidth
library(stats)
res <- optim(par=c(0.3, 0.3, 0.3), fn = CV, lower = c(1/5,1/5,1/5), upper=c(1/3,1/3,1/3), method="L-BFGS-B")
alpha_star <-res$par
alpha_star
# define es_nj(x, X_N, Y_N, h) for bandwidth CV j in {1, 2, 3}
# CV for simulation purpose => OK to diff with f_prime
# otherwise ==> diff with Y observed
# vectorize kernel estimators using outer()
# check boundary issue X has support (0, 1) and Kernel has support [-1, 1] => reflection method

# results ==> saturating at upper bound to minimize the CV function and maintain asymptotic results of Theorem 3.2
```


## Simulation Implementation

```{r}

## Dans les graphiques -> que prennent-ils comme noyaux ? => pas un choix important
## Dans les graphiques -> que fixent-ils comme bandwidth ? => validation croisée

f <- function(x) {
  return((x+2)*sin(4*pi*x^2)+2*sin(8*pi*x))
}
f_prime <- function(x) {
  return(sin(4*pi*x^2)+8*pi*x*(x+2)*cos(4*pi*x^2) + 8*pi*cos(8*pi*x))
}

# if 1/2 < p <= 1
# then B(2p-1) if 1 => I2 if 0 => I1 or I3
# if 1 => unif(I2)
# if 0 => B(1-p) => if 1 => unif(I1), else unif(I3)

# a.s. conv checked with N = 10000
# plotting es_1, es_2, es_3 against f_prime on [0, 1]
# Epanechikov Kernel

alpha <- 0.3263603
e_N <- function(x) {
  return((1/x^alpha))
}
h <- sapply(seq(1, N, 1), e_N) # bandwidth

# K values
# gaussian -> pnorm(x)
# quadratic -> ifelse(abs(x) <= 1, (15/16)*(1-x^2)^2), 0)
# epanechnikov -> ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)

# K prime values
# gaussian -> -(x*(1/sqrt(2*pi))*(exp(-(1/2)*x^2)
# quadratic -> ifelse(abs(x) <= 1, -(15/8)*(x)*(1-x^2), 0)
# epanechnikov -> ifelse(abs(x) <= 1, -(3/2)*(x), 0)


N <- 10000
eps_N <- rnorm(N, 0, 1)
X_N <- runif(N, 0, 1) # p = 1
Y_N <- f(X_N) + eps_N

g <- function(x) {
  return(1)
}
g_prime <- function(x){
  return(0)
}

K <- function(x) {
  return(ifelse(abs(x) <= 1, (3/4)*(1-x^2), 0)) 
}
K_prime <- function(x) {
  return(ifelse(abs(x) <= 1, -(3/2)*x, 0))
}

# estimator n°1
h_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
g_N <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
  }
  return(y)
}
h_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
g_N_prime <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}
es_n1 <- function(x) {
  return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
}
# estimator n°2
es_n2 <- function(x) {
  return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
}
# estimator n°3
es_n3 <- function(x) {
  y <- rep(NA, length(x))
  for (j in 1:length(x)) {
    y[j] <- (1/N)*(sum(Y_N/(g(x)*h^2)*K_prime((x[j]-X_N)/h)))
  }
  return(y)
}


# asymptotic normality checked at x = 0.5 and x = 0.8
# N = 2000

```

## Figure 1 (almost sure convergence)

```{r}
library(ggplot2)
x <- seq(0, 1, length.out = 1000)
df <- data.frame(
  x = rep(x, 4),
  y = c(f_prime(x), es_n1(x), es_n2(x), es_n3(x)),
  curve = factor(rep(c("Target","Estimator 1", "Estimator 2", "Estimator 3"), each = length(x)))
)
ggplot(df, aes(x = x, y = y, color = curve, linetype = curve)) +
  geom_line(size = 0.5) +
  geom_point(
    data = subset(df, curve != "Target"),
    size = 0.5,
    alpha = 0.7
  ) +
  scale_linetype_manual(values = c("Target" = "solid", "Estimator 1" = "dotted", "Estimator 2" = "dotted", "Estimator 3" = "dotted")) +
  labs(
    title = "Customized Curves on [0, 1]",
    x = "x",
    y = "y",
    color = "Function",
    linetype = "Function"
  ) +
  theme_minimal()

```

## Figure 2 (CLT)

```{r}
N <- 10000
B <- 200
x0 <- 0.5
x1 <- 0.8
par(mfrow=c(3,2))
set.seed(123)
ES1_X0 <- rep(NA, B)
ES2_X0 <- rep(NA, B)
ES3_X0 <- rep(NA, B)
ES1_X1 <- rep(NA, B)
ES2_X1 <- rep(NA, B)
ES3_X1 <- rep(NA, B)
for (b in 1:200) {
  e_N <- function(x) {
  return((1/x^alpha))
  }
  h <- sapply(seq(1, N, 1), e_N)
  eps_N <- rnorm(N, 0, 1)
  X_N <- runif(N, 0, 1)
  Y_N <- f(X_N) + eps_N
  h_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h)*K((x[j]-X_N)/h)))
    }
    return(y)
  }
  h_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((Y_N/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  g_N_prime <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum((1/h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  es_n1 <- function(x) {
    return((h_N_prime(x)/g_N(x)) - (h_N(x)*g_N_prime(x))/(g_N(x))^2)
    }
  es_n2 <- function(x) {
    return((h_N_prime(x)/g(x)) - (h_N(x)*g_prime(x))/(g(x))^2)
    }
  es_n3 <- function(x) {
    y <- rep(NA, length(x))
    for (j in 1:length(x)) {
      y[j] <- (1/N)*(sum(Y_N/(g(x)*h^2)*K_prime((x[j]-X_N)/h)))
    }
    return(y)
  }
  ES1_X0[b] <- sqrt(N*(1/N^alpha)^3)*
(es_n1(x0) - f_prime(x0))
  ES2_X0[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x0) - f_prime(x0))
  ES3_X0[b] <- sqrt(N*(1/N^alpha)^3)*(es_n3(x0) - f_prime(x0))
  ES1_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n1(x1) - f_prime(x1))
  ES2_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n2(x1) - f_prime(x1))
  ES3_X1[b] <- sqrt(N*(1/N^alpha)^3)*(es_n3(x1) - f_prime(x1))
}

mean0 <- mean(X0); sd0 <- sd(X0)
mean1 <- mean(X1); sd1 <- sd(X1)
x_range <- range(c(X0, X1))
x_vals <- seq(x_range[1], x_range[2], length.out = 200)
hist(X0, freq = FALSE, col = "blue",
     xlab = "Value", ylim = c(0, max(density(X0)$y, density(X1)$y)*1.2))
lines(x_vals, dnorm(x_vals, mean0, sd0), col = "red", lwd = 2)
hist(X1, freq = FALSE, col = "blue")
lines(x_vals, dnorm(x_vals, mean1, sd1), col = "red", lwd = 2)

```


## Question 3: Comparing Kernels with MSE criterion
## x = 0.5 and x = 0.8
## Question : Comparing Kernels with MISE criterion

```{r}
# values of alpha pre-defined with Q2 b)
# N <- 2000
# B <- 200
# x0 <- 0.5
# x1 <- 0.8
# MSE_K_ES = MEAN(ES(X) - f_prime(X))^2 + VAR(ES(X))
# MSE_K1_ES1(x0)
# MSE_K2_ES1(x0)
# MSE_K3_ES1(x0)
# MSE_K1_ES1(x1)
# MSE_K2_ES1(x1)
# MSE_K3_ES1(x1)

# MSE_K1_ES2(x0)
# MSE_K2_ES2(x0)
# MSE_K3_ES2(x0)
# MSE_K1_ES2(x1)
# MSE_K2_ES2(x1)
# MSE_K3_ES2(x1)

# MSE_K1_ES3(x0)
# MSE_K2_ES3(x0)
# MSE_K3_ES3(x0)
# MSE_K1_ES3(x1)
# MSE_K2_ES3(x1)
# MSE_K3_ES3(x1)

# Riemann approximation
# D <- value of discretization of [0, 1]
# SUPP <- (1:D)^(-1)
# MISE_K_ES = sum_{k=1}^D (1/D)*(MEAN(ES(x_k) - f_prime(x_k))^2 + VAR(ES(x_k))) 
# MISE_K1_ES1
# MISE_K2_ES1
# MISE_K3_ES1
# MISE_K1_ES2
# MISE_K2_ES2
# MISE_K3_ES2
# MISE_K1_ES3
# MISE_K2_ES3
# MISE_K3_ES3
```

# Annexe

## Vérification des hypothèses pour les noyaux

Ajouter les vérifications


## Calculs de l'espérance et de la variance des estimateurs


$$Y_n = f(X_n) + \epsilon_n, X_n \perp \epsilon_n, X_n \perp X_k (k \neq n), P(X_k \in B) = P(X_n \in B), B \in \mathbf{B}(\mathbf{R}), E(\epsilon) = 0, E(\epsilon^2) = \sigma^2, g \in C_b(X) $$
## Estimateur 1

$$ \left\{
    \begin{array}{ll}
       E[\hat{f}(x)] =  \\
       Var(\hat{f}(x)) =
    \end{array}
\right.$$


## Estimateur 2

$$ \left\{
    \begin{array}{ll}
       E[\tilde{f}(x)] =  \\
       Var(\tilde{f}(x)) =
    \end{array}
\right.$$


## Estimateur 3

$$ \left\{
    \begin{array}{ll}
       E[\overset{\vee}{f}(x)] = \frac{1}{n}\sum_{k=1}^n \frac{1}{h_k}E[\frac{f(X_k)}{g(X_k)}K(\frac{x-X_k}{h_k})]   \\
       
       Var(\overset{\vee}{f}(x)) =
    \end{array}
\right.$$
